---
title: "Experimental designs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Experimental designs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

`agriGLMflow` follows a design-first rule: the experimental unit and dependence structure are defined before selecting a response family. Automated choices are auditable and are not based solely on P values or the smallest information criterion.

# Core designs

The design object identifies treatments, blocks and experimental units. CRD, RCBD, Latin square and factorial designs are directly supported.

```{r}
data(agri_insects)
d <- agri_design(agri_insects, "rcbd", treatment="treatment", block="block")
agri_design_info(d)
agri_design_formula(d, "insects")
```

When blocks are random, a random intercept is generated. A user may explicitly declare a blocking factor fixed, but the choice is recorded.

# Why design precedes family

A Poisson GLM and Poisson GLMM have the same conditional response family but different independence assumptions. Family screening cannot be scientifically valid until the experimental unit is known.
