---
title: "Hierarchical agricultural designs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Hierarchical agricultural designs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

`agriGLMflow` follows a design-first rule: the experimental unit and dependence structure are defined before selecting a response family. Automated choices are auditable and are not based solely on P values or the smallest information criterion.

# Split-plot

Whole plots and subplots have distinct randomization strata.

```{r}
data(agri_splitplot)
d <- agri_design(agri_splitplot, "split_plot", block="block",
  whole_plot_factor="irrigation", subplot_factor="nitrogen",
  whole_plot_id="whole_plot_id")
d$random_terms
```

# Split-split plot

```{r}
data(agri_split_split)
d3 <- agri_design(agri_split_split, "split_split", block="block",
  whole_plot_factor="irrigation", subplot_factor="nitrogen",
  subsubplot_factor="biostimulant", whole_plot_id="whole_plot_id",
  subplot_id="subplot_id")
d3$random_terms
```

# Strip-plot

Two strip errors are represented separately. Treating a strip-plot as an ordinary factorial RCBD is explicitly avoided.

```{r}
data(agri_stripplot)
ds <- agri_design(agri_stripplot, "strip_plot", block="block",
  strip_A="tillage", strip_B="cultivar", strip_A_id="strip_A_id",
  strip_B_id="strip_B_id")
```
