---
title: "Binomial and proportional responses"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Binomial and proportional responses}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

`agriGLMflow` follows a design-first rule: the experimental unit and dependence structure are defined before selecting a response family. Automated choices are auditable and are not based solely on P values or the smallest information criterion.

# Binomial sampling

When the number of successes out of a known total is observed, use a binomial or beta-binomial model rather than treating the observed percentage as Gaussian.

```{r}
data(agri_disease)
d <- agri_design(agri_disease,"rcbd",treatment="treatment",block="block")
r <- agri_response(agri_disease,"diseased",denominator="total")
f <- agri_design_formula(d,"cbind(diseased,total-diseased)")
```

# Continuous proportions

For $0<Y<1$, beta and simplex regression are available. Closed proportions containing exact zeros or ones can use ordered beta, extended-support beta or suitable GAMLSS families.

```{r}
data(agri_cover)
agri_response(agri_cover,"cover")
```
