---
title: "Positive continuous responses"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Positive continuous responses}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

`agriGLMflow` follows a design-first rule: the experimental unit and dependence structure are defined before selecting a response family. Automated choices are auditable and are not based solely on P values or the smallest information criterion.

Strictly positive agronomic measurements may be right-skewed. Candidate models include Gamma, inverse Gaussian, lognormal, Tweedie and selected GAMLSS families.

```{r}
data(agri_biomass)
d <- agri_design(agri_biomass,"rcbd",treatment="treatment",block="block")
r <- agri_response(agri_biomass,"biomass")
agri_family_candidates(r,d)
```

The choice is evaluated with residual behavior, uncertainty and predictive adequacy, not skewness alone. A lognormal response model differs conceptually from log-transforming the observed data and fitting an ordinary Gaussian model.
