---
title: "GAMLSS and distributional regression"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{GAMLSS and distributional regression}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

`agriGLMflow` follows a design-first rule: the experimental unit and dependence structure are defined before selecting a response family. Automated choices are auditable and are not based solely on P values or the smallest information criterion.

# Beyond the mean

GAMLSS allows distribution parameters to depend on explanatory variables. A generic model can be written

$$g_\mu(\mu_i)=\eta_{\mu i},\quad g_\sigma(\sigma_i)=\eta_{\sigma i},\quad g_\nu(\nu_i)=\eta_{\nu i},\quad g_\tau(\tau_i)=\eta_{\tau i}.$$

```{r}
data(agri_distreg)
m <- agri_gamlss(data=agri_distreg, formula=biomass~dose+treatment,
  family="gamlss_GA", engine_args=list(sigma.formula=~dose))
agri_diagnose(m)
```

GAMLSS is not treated as an automatic replacement for a complex likelihood-based GLMM. When a split-plot or other hierarchical design requires random experimental-unit effects, the workflow first protects that structure.

# Expert GAMLSS families

Additional exported GAMLSS family constructors can be used explicitly through
`agri_family_spec()`. They are kept outside automatic screening until their
parameterization, diagnostics and post-hoc interpretation have dedicated tests.

```{r}
expert <- agri_family_spec("gamlss", "TF", response_type="continuous")
# agri_model(data=d, formula=y~x, family=expert, engine="gamlss")
```

# Validated and advanced GAMLSS registry

The registry separates automatic recommendations from advanced distributional
families. In addition to the conventional Gamma, lognormal and beta paths,
Tier 2 contains candidates such as Student-t, power-exponential, SHASH/JSU,
generalized Gamma, inverse Gaussian, Weibull, BCCG, BCPE, BCT, NBI/NBII,
Poisson-inverse Gaussian, generalized/double Poisson, Sichel, beta-negative
binomial, beta-binomial, simplex and zero-adjusted/inflated families when their
support matches the response.

```{r}
all_fam <- agri_families()
subset(all_fam, engine == "gamlss" & tier <= 2,
       select = c(id, response_type, multiparameter, tier))

# Explicitly allow advanced candidates in a family scan
# agri_family_scan(data=d, response="y", design=des, tier=2, fit=TRUE)
```

Tier 2 is intentionally opt-in. A family is not promoted merely because its
constructor exists in `gamlss.dist`; it must also have a defined support,
parameter interpretation, diagnostic route and reproducible tests in
`agriGLMflow`.
