---
title: "Ordinal responses"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Ordinal responses}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

`agriGLMflow` follows a design-first rule: the experimental unit and dependence structure are defined before selecting a response family. Automated choices are auditable and are not based solely on P values or the smallest information criterion.

# Ordered outcomes

Injury, disease severity and phenological stages are often ordinal rather than continuous. `agriGLMflow` supports cumulative mixed models through `ordinal` and, for fixed-effects designs, multiple VGAM formulations: cumulative, partial/non-proportional, adjacent-category, continuation-ratio and stopping-ratio models.

```{r}
data(agri_ordstage)
m1 <- agri_ordinal(agri_ordstage, stage~treatment, model="cumulative")
m2 <- agri_ordinal(agri_ordstage, stage~treatment, model="adjacent")
```

The model should reflect the scientific process. A continuation-ratio model can be meaningful when categories represent sequential progression, whereas an adjacent-category model compares neighboring stages.


## Parallelism and partial proportional odds

A proportional-odds VGAM cumulative model uses `parallel = TRUE`. Use `agri_check_parallel()` to compare it with the nested non-parallel model by likelihood ratio, while inspecting sparse categories and convergence. Partial proportional-odds models require an explicit VGAM constraint formula.

```{r, eval=FALSE}
fit_po <- agri_ordinal(agri_ordstage, stage ~ treatment, model="cumulative", parallel=TRUE)
agri_check_parallel(fit_po)
fit_ppo <- agri_ordinal(agri_ordstage, stage ~ treatment + dose,
                        model="partial_proportional",
                        parallel = TRUE ~ -1 + treatment)
```
