---
title: "Quantitative treatments and dose-response"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Quantitative treatments and dose-response}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

`agriGLMflow` follows a design-first rule: the experimental unit and dependence structure are defined before selecting a response family. Automated choices are auditable and are not based solely on P values or the smallest information criterion.

# Preserve quantitative treatments

Doses are not converted to factors automatically.

```{r}
data(agri_dose)
q <- agri_regression(agri_dose,"yield","dose",model="quadratic")
q$targets
```

Available fixed-form curves include linear, quadratic, cubic, linear plateau, quadratic plateau, Mitscherlich, Michaelis-Menten, logistic, Gompertz and Weibull models. GAM and VGAM smoothers provide flexible alternatives.

Polynomial terms can be embedded in GLMMs so a quantitative treatment can coexist with non-Gaussian distributions and random blocks. Nonlinear mixed models are not silently approximated.
