---
title: "Multiple comparisons and planned contrasts"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Multiple comparisons and planned contrasts}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

`agriGLMflow` follows a design-first rule: the experimental unit and dependence structure are defined before selecting a response family. Automated choices are auditable and are not based solely on P values or the smallest information criterion.

# Post-model inference

`emmeans` is used for model classes it supports natively. Available adjustments include Tukey, Dunnett, Holm, Sidak and FDR-compatible methods.

```{r}
fit <- agri_model(agri_insects, response="insects", design=des, family="nbinom2")
agri_means(fit,"treatment",scale="response")
agri_contrasts(fit,"treatment",method="pairwise",adjust="tukey")
```

For VGAM multinomial or compositional models, the post-hoc target is a standardized category probability or composition rather than an artificial scalar mean. `agri_means()` therefore standardizes predictions over the observed covariate distribution and can bootstrap confidence intervals.

# Treatment-versus-control comparisons

Dunnett inference is exposed explicitly rather than being treated as a synonym
for all pairwise comparisons.

```{r}
agri_contrasts(
  fit,
  specs = "treatment",
  method = "dunnett",
  control = "Control",
  scale = "response"
)
```

For model classes natively supported by `emmeans`, the package uses the
corresponding treatment-versus-control contrast family. For standardized VGAM
and GLMMadaptive predictions, a bootstrap route is used when simultaneous
inference is requested; unsupported adjustments are reported rather than
silently approximated.
