---
title: "Model diagnostics"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Model diagnostics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
# The code below is displayed rather than knitted, but R CMD check (R >= 4.6.1)
# extracts and runs it in a clean process. Attaching the package here is what
# makes that step pass and keeps the examples copy-pasteable.
library(agriGLMflow)
```

`agriGLMflow` follows a design-first rule: the experimental unit and dependence structure are defined before selecting a response family. Automated choices are auditable and are not based solely on P values or the smallest information criterion.

```{r}
data(agri_insects)
des <- agri_design(agri_insects, "crd", treatment="treatment")
fit <- agri_model(agri_insects, response="insects", design=des,
                  family="poisson", engine="stats")
```

Diagnostics are organized as gates: convergence first, then dispersion, zeros, residual behavior, random-effect boundaries, influence and residual dependence.

```{r}
dg <- agri_diagnose(fit, simulate=TRUE, nsim=500)
dg
agri_check_dispersion(fit)
agri_plot(fit,type="diagnostics")
```

A model with a smaller AIC is not promoted when its fit fails or the experimental design has been violated. Simulation-based DHARMa residuals are used when installed and supported. GAMLSS retains randomized quantile residuals, while VGAM uses native residuals, predicted category probabilities and `simulate.vlm` where implemented.
