---
title: "Multi-environment experiments"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Multi-environment experiments}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

`agriGLMflow` follows a design-first rule: the experimental unit and dependence structure are defined before selecting a response family. Automated choices are auditable and are not based solely on P values or the smallest information criterion.

# Environment is mandatory

```{r}
data(agri_multienv)
d <- agri_design(agri_multienv,"multi_environment",genotype="genotype",
  environment="environment",block="block")
```

A multi-environment declaration without `environment` is an error. Blocks are nested within environments, and genotype, environment and genotype-by-environment terms may be declared fixed or random.

```{r}
# Leakage-safe validation
# cv <- agri_cv(fit, scheme="leave_one_environment_out")
```
