---
title: "Scientific graphics and reporting"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Scientific graphics and reporting}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

`agriGLMflow` follows a design-first rule: the experimental unit and dependence structure are defined before selecting a response family. Automated choices are auditable and are not based solely on P values or the smallest information criterion.

`agriGLMflow` graphics return `ggplot2` objects. The default philosophy is observed data plus model estimates and uncertainty rather than isolated mean bars.

```{r}
agri_plot_data(agri_insects,"insects","treatment",type="boxplot")
agri_plot(fit,type="fit")
agri_plot(fit,type="zero")
agri_plot(fit,type="means",specs="treatment")
```

Reports record design, response type, family, engine, convergence, diagnostics and the audit trail. Markdown requires no external document converter; HTML, DOCX and PDF can be rendered with `rmarkdown` when available.
