---
title: "Simulation, bootstrap and power"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Simulation bootstrap and power}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

`agriGLMflow` follows a design-first rule: the experimental unit and dependence structure are defined before selecting a response family. Automated choices are auditable and are not based solely on P values or the smallest information criterion.

# Simulation

```{r}
sim <- agri_simulate(fit,nsim=100,seed=2026)
boot <- agri_bootstrap(fit,R=500,type="cluster",unit="block")
pow <- agri_power(fit,term="treatment",nsim=1000)
```

Cluster bootstrap resamples experimental units rather than assuming individual observations are exchangeable. Parametric bootstrap simulates from the fitted model. Power is simulation-based and therefore conditional on the fitted parameterization.

For software-validation studies, seeds and scenario definitions should be frozen before final package comparisons.

# Frozen validation battery A-L

The package ships a manuscript-oriented simulation plan under
`inst/simulation/`. The frozen scenario grid covers count-family selection,
proportions, positive continuous responses, GAMLSS location-scale models,
design misspecification, quantitative regression, multiplicity, diagnostic
sensitivity, ordinal VGAM models, multinomial VGAM models, compositional VGAM
models and cross-backend parameterization checks.

The validation endpoints include coefficient bias, RMSE, interval coverage,
type-I error, power, convergence, family-set recovery, predictive log score,
Brier score, multiclass log-loss, calibration and computational success. The
scenario definitions and seed registry are versioned so that later package
changes cannot silently alter the article experiments.

```{r}
# Installed-package path
simdir <- system.file("simulation", package = "agriGLMflow")
read.csv(file.path(simdir, "scenario_grid.csv")) |> head()
```
