---
title: "Advanced engines and routing"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Advanced engines}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

`agriGLMflow` follows a design-first rule: the experimental unit and dependence structure are defined before selecting a response family. Automated choices are auditable and are not based solely on P values or the smallest information criterion.

# Engine roles

* `stats`: conventional fixed-effects GLM.
* `glmmTMB`: primary advanced GLMM engine.
* `gamlss`: distributional regression.
* `VGAM`: fixed-effects vector, multinomial, ordinal and compositional models.
* `lme4`: reference/fallback engine for conventional GLMMs and Gaussian mixed models.
* `GLMMadaptive`: adaptive Gauss-Hermite quadrature for compatible models with one grouping factor.
* `ordinal`: cumulative link mixed models.
* `betareg`: beta and extended-support beta regression.
* `brglm2`: bias-reduced GLM when separation or small-sample bias is relevant.

```{r}
agri_engines()
agri_dependencies()
```

An explicit engine request never authorizes the package to drop required random effects.
