---
title: "Vector, multinomial and compositional models with VGAM"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{VGAM integration}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

`agriGLMflow` follows a design-first rule: the experimental unit and dependence structure are defined before selecting a response family. Automated choices are auditable and are not based solely on P values or the smallest information criterion.

# Why VGAM is included

VGAM extends the package beyond ordinary scalar-response GLMs. It provides multinomial, cumulative and alternative ordinal models, Dirichlet and Dirichlet-multinomial responses, simplex regression, specialized count distributions, censoring and extreme-value families. Current VGAM releases implement fixed-effects models only, so the engine is deliberately blocked for random-effects designs.

# Multinomial

```{r}
data(agri_multiclass)
m <- agri_multinomial(agri_multiclass,status~treatment)
agri_means(m,"treatment",bootstrap=200)
```

# Composition

```{r}
data(agri_composition)
dm <- agri_composition(agri_composition,cbind(root,stem,leaf)~treatment)
agri_means(dm,"treatment",bootstrap=200)
```

# Ordinal alternatives

```{r}
data(agri_ordstage)
cum <- agri_ordinal(agri_ordstage,stage~treatment,model="cumulative")
adj <- agri_ordinal(agri_ordstage,stage~treatment,model="adjacent")
con <- agri_ordinal(agri_ordstage,stage~treatment,model="continuation")
```

# Capability tiers

Tier 1 families enter the automated candidate registry after dedicated tests. Tier 2 contains advanced or numerically fragile models. Wider VGAM families can be accessed only by explicit expert-level extension until their post-hoc and diagnostic semantics are validated. This prevents the phrase “100+ distributions” from becoming an unvalidated automatic-search space.

# Expert access to additional VGAM families

Families not yet in the validated registry remain accessible without pretending that
they passed automatic screening:

```{r}
expert <- agri_family_spec("VGAM", "some_exported_VGAM_family",
  response_type="expert", posthoc_mode="unsupported")
# agri_model(data=d, formula=y~x, family=expert, engine="VGAM")
```

Expert specifications never enter `agri_family_scan()` automatically.

# Dirichlet-multinomial counts

```{r}
data(agri_multicounts)
dmc <- agri_composition(
  agri_multicounts,
  cbind(healthy,mild,moderate,severe) ~ treatment,
  family="dirmultinomial"
)
```
