---
title: "agriGLMflow Tutorial: Complete Modeling Workflows for Agronomic Experiments"
subtitle: "From Planning to Inference in English"
author: "Walter E. Pereira and Magali H. P. Martinez"
date: today
format:
  html:
    toc: true
    number-sections: true
    code-fold: show
    embed-resources: true
    theme: cosmo
execute:
  echo: true
  warning: false
  message: false
  eval: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 5,
  fig.align = "center"
)
set.seed(2026)
# Load the package via devtools (during development)
if (requireNamespace("agriGLMflow", quietly = TRUE)) {
  library(agriGLMflow)
} else {
  devtools::load_all(system.file(package = "agriGLMflow"), quiet = TRUE)
}
```

## Introduction {.unnumbered}

This tutorial shows, step by step, how to use the `agriGLMflow` package to
analyze agronomic experiments. Each section begins with a **domain question**,
shows the data before the formulas, and guides the complete analysis
through to inference and plots.

**What you will learn:**

- How to correctly declare the experimental design
- How to identify the appropriate distribution family
- How to fit models and compare statistical engines
- How to perform post-model inference (means, contrasts, trends)
- How to generate publication-quality graphs

```{r}
packageVersion("agriGLMflow")
R.version.string
```

## Workflow 1: Qualitative Factor (RCBD with Counts) {.unnumbered}

### Domain Question

> Does a biofertilizer reduce insect infestation in a bean crop?
> Is the effect consistent across blocks?

### 1.1 Prepare the Data

```{r}
# Load insect count data
data("agri_insects", package = "agriGLMflow")

# First rows
head(agri_insects, 10)

# Summary by treatment
aggregate(insects ~ treatment, data = agri_insects,
          FUN = function(x) c(mean = mean(x), sd = sd(x), n = length(x)))
```

### 1.2 Declare the Design

```{r}
# Create RCBD design object
des <- agri_design(
  data = agri_insects,
  design = "rcbd",
  treatment = treatment,
  block = block
)
print(des)
```

### 1.3 Characterize the Response

```{r}
# Characterize the response variable (counts)
resp <- agri_response(
  data = agri_insects,
  response = "insects"
)
print(resp)
```

### 1.4 Identify Candidate Families

```{r}
# Which families are suitable for non-negative integer counts?
cands <- agri_family_candidates(resp, tier = 1L)
print(cands)
```

### 1.5 Fit the Model

```{r}
# Fit model with auto family (automatic selection)
# The resp object (agri_response) is passed directly
fit <- agri_model(
  data = agri_insects,
  response = resp,
  design = des,
  family = "auto"
)

# Fit summary
summary(fit$engine_fit)
```

### 1.6 Diagnostics

```{r}
# Complete diagnostics
diag <- agri_diagnose(fit)
print(diag)
```

### 1.7 Post-Model Inference

```{r}
# Marginal means by treatment
emm <- agri_means(fit, specs = ~ treatment, scale = "response")
print(emm)
```

```{r}
# Tukey contrasts (pairwise treatment comparisons)
# Note: for glmmTMB Poisson models with block, contrasts
# may require explicit specs
contr <- tryCatch(
  agri_contrasts(fit, specs = ~ treatment, method = "pairwise",
                 adjust = "tukey", scale = "response"),
  error = function(e) {
    cat("Contrasts via emmeans directly:\n")
    emmeans::emmeans(fit$engine_fit, pairwise ~ treatment, type = "response")
  }
)
print(contr)
```

### 1.8 Plots

```{r}
#| fig-cap: "Marginal means by treatment with confidence intervals"
# Means plot with ggplot2 directly
library(ggplot2)
emm_df <- emm$table
ggplot(emm_df, aes(x = treatment, y = response)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = asymp.LCL, ymax = asymp.UCL), width = 0.2) +
  labs(x = "Treatment", y = "Mean Insects",
       title = "Marginal Means by Treatment") +
  theme_minimal()
```

```{r}
#| fig-cap: "Residual diagnostics"
# Residual plot with ggplot2
# For glmmTMB, use fitted() and residuals()
resid_df <- data.frame(
  fitted = fitted(fit$engine_fit),
  residuals = residuals(fit$engine_fit, type = "pearson")
)
ggplot(resid_df, aes(x = fitted, y = residuals)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_smooth(se = FALSE, color = "red") +
  labs(x = "Fitted Values", y = "Pearson Residuals",
       title = "Residuals vs Fitted Values") +
  theme_minimal()
```

### Lesson

> For counts with a block effect, the Poisson model with a random block
> preserves the experimental unit and provides estimates with statistical
> validity. Do not use `aov()` for counts: it assumes normality.

## Workflow 2: Quantitative Factor (Dose-Response Regression) {.unnumbered}

### Domain Question

> What is the optimal nitrogen dose that maximizes yield?
> Does the response follow a quadratic curve with a plateau?

### 2.1 Prepare the Data

```{r}
data("agri_dose", package = "agriGLMflow")
head(agri_dose)

# Exploratory plot
plot(yield ~ dose, data = agri_dose,
     main = "Yield vs Nitrogen Dose",
     xlab = "Dose (kg/ha)", ylab = "Yield (t/ha)")
```

### 2.2 Fit Regression Models

```{r}
# Quadratic model
fit_quad <- agri_regression(
  data = agri_dose,
  response = "yield",
  x = "dose",
  model = "quadratic"
)

# Summary with parameters and agronomic targets
print(fit_quad)
```

### 2.3 Compare Models

```{r}
# Also fit a linear-plateau model
fit_lp <- agri_regression(
  data = agri_dose,
  response = "yield",
  x = "dose",
  model = "linear_plateau"
)

# Compare goodness of fit
# For nls models, check summary structure
cat("Quadratic - fit summary:\n")
s_q <- summary(fit_quad$fit)
cat("  R²:", round(s_q$r.squared %||% NA, 4), "\n")
cat("  RMSE:", round(s_q$sigma %||% NA, 4), "\n")

cat("\nLinear-Plateau - summary:\n")
s_lp <- summary(fit_lp$fit)
cat("  Parameters:\n")
print(coef(s_lp))

# Show quadratic model parameters
cat("\nQuadratic parameters:\n")
print(coef(fit_quad$fit))
```

### 2.4 Regression Plot

```{r}
#| fig-cap: "Quadratic regression curve"
# Regression plot with ggplot2
newdata <- data.frame(dose = seq(min(agri_dose$dose), max(agri_dose$dose), length.out = 100))
pred <- predict(fit_quad$fit, newdata = newdata, se.fit = TRUE)
newdata$yield_pred <- pred$fit
newdata$yield_se <- pred$se.fit

ggplot(agri_dose, aes(x = dose, y = yield)) +
  geom_point() +
  geom_line(data = newdata, aes(x = dose, y = yield_pred), color = "blue") +
  geom_ribbon(data = newdata, aes(x = dose, ymin = yield_pred - 1.96*yield_se,
                                   ymax = yield_pred + 1.96*yield_se),
              inherit.aes = FALSE, alpha = 0.2) +
  labs(x = "Dose (kg/ha)", y = "Yield (t/ha)",
       title = "Quadratic Dose-Response Regression") +
  theme_minimal()
```

### Lesson

> Variables such as dose, time, or rate remain quantitative by default.
> Do not silently convert them to factors: this loses information and
> statistical power.

## Workflow 3: Comparing Statistical Engines {.unnumbered}

### Domain Question

> For soil cover proportion data, which statistical engine produces the
> best fit? Betareg, GAMLSS, or glmmTMB?

### 3.1 Prepare the Data

```{r}
data("agri_cover", package = "agriGLMflow")
head(agri_cover)

# Check proportion limits
range(agri_cover$cover)
```

### 3.2 Fit with Multiple Engines

```{r}
# Betareg
fit_br <- tryCatch(
  agri_model(data = agri_cover, response = "cover",
             formula = cover ~ treatment, family = "extended_beta",
             engine = "betareg"),
  error = function(e) NULL
)

# glmmTMB beta
fit_tm <- tryCatch(
  agri_model(data = agri_cover, response = "cover",
             formula = cover ~ treatment, family = "beta",
             engine = "glmmTMB"),
  error = function(e) NULL
)

# GAMLSS beta
fit_ga <- tryCatch(
  agri_model(data = agri_cover, response = "cover",
             formula = cover ~ treatment, family = "BEINF",
             engine = "gamlss"),
  error = function(e) NULL
)

# Check which fits were successful
cat("betareg:", !is.null(fit_br), "\n")
cat("glmmTMB:", !is.null(fit_tm), "\n")
cat("gamlss:", !is.null(fit_ga), "\n")
```

### 3.3 Comparative Metrics Table

```{r}
# Build comparison table
metrics <- data.frame(
  Engine = character(),
  AIC = numeric(),
  logLik = numeric(),
  stringsAsFactors = FALSE
)

if (!is.null(fit_br)) {
  metrics <- rbind(metrics, data.frame(
    Engine = "betareg", AIC = AIC(fit_br$engine_fit),
    logLik = as.numeric(logLik(fit_br$engine_fit))
  ))
}
if (!is.null(fit_tm)) {
  metrics <- rbind(metrics, data.frame(
    Engine = "glmmTMB", AIC = AIC(fit_tm$engine_fit),
    logLik = as.numeric(logLik(fit_tm$engine_fit))
  ))
}
if (!is.null(fit_ga)) {
  metrics <- rbind(metrics, data.frame(
    Engine = "gamlss", AIC = AIC(fit_ga$engine_fit),
    logLik = as.numeric(logLik(fit_ga$engine_fit))
  ))
}

metrics <- metrics[order(metrics$AIC), ]
print(metrics)
```

### Lesson

> The metrics table is **honest**: it includes all engines tested,
> even when one stands out. A very different AIC may indicate different
> parameterization, not necessarily a better fit. Always check the
> parameterization of each backend before comparing.

## Workflow 4: Diagnostics and Sensitivity {.unnumbered}

### Domain Question

> Is the count model adequate? Is there overdispersion or excess zeros?

### 4.1 Detailed Diagnostics

```{r}
# Use the fit from Workflow 1
diag <- agri_diagnose(fit, simulate = TRUE, nsim = 100)

# Overall status
cat("Overall status:", diag$status, "\n")

# Diagnostic components
names(diag)
```

### 4.2 Sensitivity to Alternative Families

```{r}
# Test alternative families to check robustness
# Note: agri_sensitivity may fail with certain combinations
sens <- tryCatch(
  agri_sensitivity(fit, families = c("poisson", "nbinom2", "nbinom1")),
  error = function(e) {
    cat("Sensitivity not available:", conditionMessage(e), "\n")
    NULL
  }
)
if (!is.null(sens)) print(sens)
```

### 4.3 Cross-Validation

```{r}
# 5-fold cross-validation respecting blocks
cv_result <- agri_cv(fit, v = 5L, unit = "block")
print(cv_result)
```

### Lesson

> An acceptable diagnostic does not guarantee the best model; it ensures
> that the model has no evident problems. Always compare with alternative
> families using prediction metrics.

## Closing {.unnumbered}

### Complete Package Workflow

```{mermaid}
graph TD
    A[Raw data] --> B[agri_design]
    B --> C[agri_response]
    C --> D[agri_family_scan]
    D --> E[agri_model]
    E --> F[agri_diagnose]
    F --> G[agri_means / agri_contrasts]
    G --> H[agri_plot_*]
    E --> I[agri_regression]
    I --> J[agri_compare_curves]
    F --> K[agri_cv / agri_bootstrap]
    E --> L[agri_report]
```

### Best Practice Rules

1. **Declare the design before fitting**: the experimental unit
   determines the variance structure.
2. **Do not p-hack**: model selection should be based on predefined
   criteria (AIC, BIC, cross-validation), not on significance.
3. **Always check residuals**: convergence is not sufficient.
4. **Record decisions**: use `agri_audit()` to track what was done
   and why.
5. **Use fixed seeds**: reproducibility is essential for science.

### Exercises

<details>
<summary>Exercise 1 (Basic Level)</summary>

Fit a Poisson model to the `agri_disease` data using `agri_glm()`.
Compare the results with a direct call to `glm()`.

```{r}
# Your answer here
```

</details>

<details>
<summary>Exercise 2 (Intermediate Level)</summary>

For the `agri_germination` data, fit a binomial model with `agri_glm()`
and another with `agri_model(family = "auto")`. Compare the estimates.

```{r}
# Your answer here
```

</details>

<details>
<summary>Exercise 3 (Advanced Level)</summary>

Use `agri_workflow()` to analyze the `agri_splitplot` data as a
split-plot design. Verify that the package correctly detected the
main-plot structure.

```{r}
# Your answer here
```

</details>

### How to Cite

```{r}
citation("agriGLMflow")
```

```{r}
packageVersion("agriGLMflow")
R.version.string
```

```{r}
sessionInfo()
```
